### 2016-07-17 Chapters/Chapter02.tex
Add notation chapter

%*****************************************
\chapter{Response Theory and Molecular Properties}\label{ch:molprop}
%*****************************************
\epigraph{What's the frequency, Kenneth?}{--- \textsc{R.E.M.}}

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************