### 2016-07-17 Chapters/Ch01-ContinuumSolvationModels.tex
Start on papers summaries and contributions

%************************************************
\chapter{Continuum solvation models}\label{ch:CSM}
%************************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************