### 2016-07-17 Chapters/Ch02-ResponseTheory.tex

%*****************************************
\chapter{Response theory and molecular properties}\label{ch:molprop}
%*****************************************
\epigraph{What's the frequency, Kenneth?}{--- \textsc{R.E.M.}}

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************