### 2016-07-17 Chapters/PaperSummary.tex

%************************************************
\chapter{Summary of the papers}\label{ch:papers-summary}

The codes used in the papers are all interfaced to the \PCMSolver library,
of which I am the main developer and maintainer.

\section{Paper I. Continuum solvation in the relativistic regime}

I contributed the theoretical derivation of the quantum/classical polarizable
terms in a four-component \acs{SCF} framework, for energies and linear response
properties. I devised the coupling of the four-component program \DIRAC with
\PCMSolver by providingi the implementation and testing of:
\begin{itemize}
  \item \acs{MEP} integrals for four-component wave functions,
  \item the additional Fock matrix contributions,
  \item the additional terms in the response equations.
\end{itemize}
I performed all the calculations and large part of the data analysis
for the results reported in the paper.
Finally, I wrote the first draft of the paper and coordinated editing of
all subsequent versions.

\section{Paper II. The wavelet boundary element method for \acs{PCM}}
\section{Paper III: Continuum modelling of non homogeneous environments}

%************************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************