### 2016-07-17 acronyms.tex

\begin{acronym}
\acro{CC}{coupled cluster}
\acro{CCSD}{coupled cluster with single and double substitutions}
\acro{CI}{configuration interaction}
\acro{MO}{molecular orbital}
\acro{LR}{linear response}
\acro{DFT}{density-functional theory}
\acro{PCM}{polarizable continuum model}
\acro{MM}{molecular mechanics}
\acro{MP}{M{\o}ller--Plesset}
\acro{HF}{Hartree--Fock}
\acro{ASC}{apparent surface charge}
\acro{PTE}{perturbation to the energy}
\acro{PTED}{perturbation to the energy and the density}
\acro{QM/MM}{Quantum Mechanics/Molecular Mechanics}
\acro{DC}{Dielectric Continuum}
\acro{GEPOL}{GEPOL}
\acro{IEF}{Integral Equation Formalism}
\acro{LHS}{left-hand side}
\acro{MD}{Molecular Dynamics}
\acro{MM}{Molecular Mechanics}
\acro{BVP}{boundary value problem}
\acro{PDE}{partial differential equation}
\acro{PE}{Polarizable Embedding}
\acro{QC}{Quantum Chemistry}
\acro{RHS}{right-hand side}
\acro{SAS}{Solvent-Accessible Surface}
\acro{SES}{Solvent-Excluded Surface}
\acro{MEP}{molecular electrostatic potential}
\acro{TDSCF}{time-dependent SCF}
\acro{KS}{Kohn--Sham}
\end{acronym}