### 2016-07-27 Chapters/Ch02-ResponseTheory.tex
Add commands, fix notations

%*****************************************
\chapter{Response theory and molecular properties}\label{ch:molprop}
%*****************************************
\epigraph{What's the frequency, Kenneth? [...]

          I never understood the frequency, uh-huh}{
          --- \textsc{R.E.M.}}

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************