### 2016-07-27 Chapters/Ch01-ContinuumSolvationModels.tex
Some more work

% mainfile: ../RobertoDiRemigioPhDThesis.tex
\chapter{Continuum solvation models}\label{ch:CSM}

\section{A unifying view of classical polarizable models}

Variational functional for implicit models, \acs{PCM} in this case:
\begin{equation}
 U_\mathrm{PCM} = \frac{1}{2}\scalprod{\sigma}{\PCM\sigma} + \scalprod{\esp}{\sigma}
\end{equation}

Variational functional for explicit models, could be MMpol or PE or FQ:
\begin{equation}
  U_\mathrm{MM} = \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
\end{equation}

The PCM equations will be written in the ``complete basis'' meaning that
we will introduce the usual boundary-element method (BEM) discretization
at the very end of the derivation. In other words, we will be working
with the exact integral equation and not with its discretized
counterpart. As a consequence, the apparent surface charge $\sigma$ and
the electrostatic potential $\esp$ will have a \emph{continuous}
dependence on a ``cavity surface'' index $\vect{s}$. Whenever a
charge-potential product is present it is then to be interpreted as the
\emph{surface integral}, i.e. the scalar product in the suitable,
infinite-dimensional vector space on the cavity boundary $\Gamma$. The
following shorthand notation will be adopted:
\begin{equation}
 \sigma\esp = \scalprod{\sigma}{\esp}
\end{equation}

Coupled implicit/explicit polarization energy functional:
\begin{equation}
  U_\mathrm{pol} =
   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
 + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
 + \sigma\bi{X}\kappa
\end{equation}
where one has:
\begin{equation}
  \sigma\MMPCM\kappa = \kappa\MMPCM^\dagger\sigma
\end{equation}
and $\MMPCM$ is the implicit/explicit interaction kernel. This is
charge/dipole or charge/charge interaction kernel for the MMpol and PE models
or the FQ model, respectively.
The global minimum of the convex functional is found by differentiating
it with respect to the variational degrees of freedom, \ie{} the
\acl{ASC} (\acs{ASC}) density $\sigma$ and the classical point
charges/dipoles $\kappa$. This leads to the coupled equations:
\begin{align}
  \PCM\sigma + \esp + \MMPCM\kappa &= 0 \\
  \MM\kappa  + \zeta + \MMPCM^\dagger\sigma &= 0
\end{align}
or in the commonly found matrix form:
\begin{equation}
  \begin{pmatrix}
    \PCM & \MMPCM^\dagger \\
    \MMPCM & \MM
  \end{pmatrix}
  \begin{pmatrix}
   \sigma \\
   \kappa
  \end{pmatrix}
  =
  -
  \begin{pmatrix}
   \esp \\
   \zeta
  \end{pmatrix}
\end{equation}
Finally, let us re-express the equations above in a "supermatrix"
formalism:
\begin{equation}
  U_\mathrm{pol} =
  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\s
\end{equation}
where:
\begin{alignat}{3}
  \p =
  \begin{pmatrix}
    \sigma \\
    \kappa
  \end{pmatrix},
  \quad&
  \s =
  \begin{pmatrix}
   \esp \\
   \zeta
  \end{pmatrix},
  \quad&
  \V =
  \begin{pmatrix}
    \PCM & \MMPCM^\dagger \\
    \MMPCM & \MM
  \end{pmatrix}
\end{alignat}
The supermatrix form will prove particularly useful in the following
derivations.