### 2016-08-10 Chapters/Ch02-ResponseTheory.tex
Write summary for paper III

%*****************************************
\chapter{Response theory and molecular properties}\label{ch:molprop}
%*****************************************

\epigraph{
White light, white light goin' messin' up my brain

White light, aww white light it's gonna drive me insane
}{
--- \textsc{The Velvet Underground}
}

\epigraph{What's the frequency, Kenneth? [...]

          I never understood the frequency, uh-huh}{
          --- \textsc{R.E.M.}}

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************