### 2016-08-24 FrontBackmatter/Abstract.tex
Clean up and add one chapter (to be written)

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

\vfill

\endgroup

\vfill

\begin{flushright}
    \begin{tabular}{m{5cm}}
      \textit{\myLocation, \myTime} \\
      \includegraphics[width=0.30\textwidth]{gfx/signature.png} \\
        \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}