### 2016-08-26 acronyms.tex

\begin{acronym}
\acro{CC}{coupled cluster}
\acro{CCS}{coupled cluster with single substitutions}
\acro{CCSD}{coupled cluster with single and double substitutions}
\acro{CCSDT}{coupled cluster with single, double and triple substitutions}
\acro{CC2}{approximate coupled cluster singles and doubles}
\acro{CC3}{approximate coupled cluster singles, doubles and triples}
\acro{PT}{perturbation theory}
\acro{MBPT}{many-body perturbation theory}
\acro{CI}{configuration interaction}
\acro{MO}{molecular orbital}
\acro{AO}{atomic orbital}
\acro{LR}{linear response}
\acro{DFT}{density-functional theory}
\acro{PCM}{polarizable continuum model}
\acro{MM}{molecular mechanics}
\acro{MP}{M{\o}ller--Plesset}
\acro{HF}{Hartree--Fock}
\acro{ASC}{apparent surface charge}
\acro{PTE}{perturbation-to-the-energy}
\acro{PTES}{perturbation to the energy}
\acro{PTE(S)}{perturbation to the energy}
\acro{PTED}{perturbation-to-the-energy-and-the-density}
\acro{QM/MM}{Quantum Mechanics/Molecular Mechanics}
\acro{DC}{Dielectric Continuum}
\acro{GEPOL}{GEPOL}
\acro{IEF}{Integral Equation Formalism}
\acro{LHS}{left-hand side}
\acro{MD}{Molecular Dynamics}
\acro{MM}{Molecular Mechanics}
\acro{BVP}{boundary value problem}
\acro{BEM}{boundary element method}
\acro{BIE}{boundary integral equation}
\acro{PDE}{partial differential equation}
\acrodefplural{PDE}{partial differential equations}
\acro{DtN}{Dirichlet-to-Neumann}
\acro{PE}{polarizable embedding}
\acro{FQ}{fluctuating charges}
\acro{QC}{Quantum Chemistry}
\acro{RHS}{right-hand side}
\acro{SAS}{Solvent-Accessible Surface}
\acro{SES}{Solvent-Excluded Surface}
\acro{MEP}{molecular electrostatic potential}
\acro{SCF}{self-consistent field}
\acro{TDSCF}{time-dependent SCF}
\acro{KS}{Kohn--Sham}
\acro{PWC}{piecewise constant wavelet}
\acro{PWL}{piecewise constant linear}
\acro{DC}{Dirac--Coulomb}
\acro{BCH}{Baker--Campbell--Hausdorff}
\end{acronym}