### 2016-08-26 FrontBackmatter/Abstract.tex
Write some of the abstract

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

The vast majority of chemical phenomena happens in complex environments
where the molecule of interest can interact with a large number of other
moieties, be they solvent molecules or a structured environment, such as
a protein.
Spectroscopic experiments measure the response of a chemical system to
an applied perturbing electromagnetic field and can probe the
microscopic motions of electrons and nuclei in molecules.
This achieves unique insight into the structural and physical behaviour
of complex systems.
Characterizing, explaining and predicting the effect of the chemical
environment on a large number of measurable properties requires,
however, a synergistic experimental and theoretical approach.

In this thesis, I will present our efforts in the development of the
polarizable continuum model for solvation.

\vfill

% Anything in Norwegian?
%\pdfbookmark[1]{Zusammenfassung}{Zusammenfassung}
%\chapter*{Zusammenfassung}
%Kurze Zusammenfassung des Inhaltes in deutscher Sprache\dots

\endgroup

\vfill

\begin{flushright}
    \begin{tabular}{m{5cm}}
      \textit{\myLocation, \myTime} \\
      \includegraphics[width=0.40\textwidth]{gfx/signature.png} \\
        \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}