### 2016-09-01 acronyms.tex

\newacronym{API}{API}{application programming interface}
\newacronym{COSMO}{COSMO}{conductor-like screening model}
\newacronym{BO}{BO}{Born--Oppenheimer}
\newacronym[plural=PES, firstplural={potential energy surfaces (PESs)}]{PES}{PES}{potential energy surface}
\newacronym{CI}{CI}{configuration interaction}
\newacronym{CC}{CC}{coupled cluster}
\newacronym{CCS}{CCS}{coupled cluster with single substitutions}
\newacronym{CCSD}{CCSD}{coupled cluster with single and double substitutions}
\newacronym{CCSDT}{CCSDT}{coupled cluster with single, double and triple substitutions}
\newacronym{CC2}{CC2}{approximate coupled cluster singles and doubles}
\newacronym{CC3}{CC3}{approximate coupled cluster singles, doubles and triples}
\newacronym{PT}{PT}{perturbation theory}
\newacronym{MBPT}{MBPT}{many-body perturbation theory}
\newacronym{MO}{MO}{molecular orbital}
\newacronym{AO}{AO}{atomic orbital}
\newacronym{LR}{LR}{linear response}
\newacronym{DFT}{DFT}{density-functional theory}
\newacronym{PCM}{PCM}{polarizable continuum model}
\newacronym{MM}{MM}{molecular mechanics}
\newacronym{MP}{MP}{M{\o}ller--Plesset}
\newacronym{HF}{HF}{Hartree--Fock}
\newacronym{ASC}{ASC}{apparent surface charge}
\newacronym{PTE}{PTE}{perturbation-to-the-energy}
\newacronym{PTES}{PTES}{perturbation to the energy}
\newacronym{PTE(S)}{PTE(S)}{perturbation to the energy}
\newacronym{PTED}{PTED}{perturbation-to-the-energy-and-the-density}
\newacronym{IEF}{IEF}{integral equation formalism}
\newacronym{LHS}{LHS}{left-hand side}
\newacronym{MD}{MD}{molecular dynamics}
\newacronym{QM}{QM}{quantum mechanics}
\newacronym{BVP}{BVP}{boundary value problem}
\newacronym{BEM}{BEM}{boundary element method}
\newacronym{BIE}{BIE}{boundary integral equation}
\newacronym[plural=PDE, firstplural={partial differential equations (PDEs)}]{PDE}{PDE}{partial differential equation}
\newacronym{DtN}{DtN}{Dirichlet-to-Neumann}
\newacronym{PE}{PE}{polarizable embedding}
\newacronym{FQ}{FQ}{fluctuating charges}
\newacronym{QC}{QC}{quantum chemistry}
\newacronym{RHS}{RHS}{right-hand side}
\newacronym{vdWS}{vdWS}{van der Waals surface}
\newacronym{SAS}{SAS}{solvent-accessible surface}
\newacronym{SES}{SES}{solvent-excluded surface}
\newacronym{MEP}{MEP}{molecular electrostatic potential}
\newacronym{SCF}{SCF}{self-consistent field}
\newacronym{TDSCF}{TDSCF}{time-dependent SCF}
\newacronym{KS}{KS}{Kohn--Sham}
\newacronym{PWC}{PWC}{piecewise constant wavelet}
\newacronym{PWL}{PWL}{piecewise constant linear}
\newacronym{DC}{DC}{Dirac--Coulomb}
\newacronym{BCH}{BCH}{Baker--Campbell--Hausdorff}
\newacronym[plural=DVCS, firstplural={distributed version control systems (DVCSs)}]{DVCS}{DVCS}{distributed version control system}
\newacronym{EPR}{EPR}{electron paramagnetic resonance}
\newacronym{NMR}{NMR}{nuclear magnetic resonance}
\newacronym{pNMR}{pNMR}{paramagnetic nuclear magnetic resonance}
\newacronym{LGPL}{LGPL}{GNU Lesser General Public License}
\newacronym{AD}{AD}{automatic differentiation}
\newacronym{CSC}{CSC}{continuous surface charge}
\newacronym{SWIG}{SWIG}{switching/Gaussian}