### 2016-09-07 Chapters/ResponseTheory.tex
Finish introduction. I'll have to add some stuff later..:

\chapter{Response Theory and Molecular Properties}\label{ch:molprop}

\begin{epigraphs}
\qitem{
White light goin' messin' up my brain

White light it's gonna drive me insane
}{
--- \textsc{The Velvet Underground}
}
\qitem{What's the frequency, Kenneth? [...]

          I never understood the frequency, uh-huh}{
          --- \textsc{R.E.M.}}
\end{epigraphs}

Here and in the following, "external" will be used to mean an
electromagnetic field which is "weak" when compared to the
electron--electron and electron--nuclei interactions in the molecule.

Section \ref{sec:exact-response} will give a brief introduction to exact
state response theory for isolated molecules. The concepts of
quasienergy,\autocite{Christiansen1998-pe} variational perturbation
theory\autocite{Helgaker1992-ph} and pole-and-residue
analysis of response functions\autocite{Olsen1985-nr} will be introduced.
I will explicitly derive the \acrshort{SCF} response function for a
quantum/classical polarizable Hamiltonian in Section
\ref{sec:csm-response}. The derivation will employ the
density matrix-based, \acrshort{AO} formalism introduced by
\citeauthor{Thorvaldsen2008-sg} in \noparcite[ref.][]{Thorvaldsen2008-sg}.
This formalism is suitable for the derivation of arbitrary-order
response functions and their residues, also when perturbation-dependent
basis sets are considered.\autocite{Friese2015-kb, Friese2015-bu}
\paper{V} presents the arbitrary-order derivation of the \acrshort{PCM}.

The response treatment of molecular properties has its roots in
time-dependent perturbation theory\autocite{Konishi2009-zb} and has been
continually developed in quantum chemistry for the past 30 years:
the exposition will be necessarily brief and incomplete, the reader is
referred to the cited works in the literature for a more detailed
presentation of the topic.\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}

\section{Exact State Response Theory}\label{sec:exact-response}

General theory~\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
Solution strategy for response equations~\autocite{Saad2003-oa,
Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}
Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
Kristensen2008-hv, Ringholm2014-gx}
Open-ended \acrshort{SCF} residues~\autocite{Friese2015-kb}

\section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}

\acrshort{PCM}~\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
\acrshort{FQ} and \acrshort{FQ}/\acrshort{PCM}~\autocite{Lipparini2012-hx, Lipparini2012-tl, Lipparini2013-ud}
\acrshort{PE}~\autocite{Olsen2010-wa}
\acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Steindal2011-ki, Caprasecca2012-ir}
Paper V~\autocite{pcm-openrsp}