### 2016-09-27 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

\begin{epigraphs}
\setlength\epigraphwidth{6.5cm}
\qitem{
Just a perfect day

You made me forget myself

I thought I was

Someone else, someone good}{
--- \textsc{Lou Reed}}
\qitem{
And all of my friends were there,

and no one cared}{
--- \textsc{The Kinks}}
\setlength\epigraphwidth{7cm}
\end{epigraphs}

\endgroup