### 2016-09-28 FrontBackmatter/Colophon.tex
Switch to Minion Pro

\pagestyle{empty}

\hfill

\vfill

\pdfbookmark[1]{Colophon}{colophon}
\section*{Colophon}
This document was typeset using \XeLaTeX, with the
\texttt{classicthesis} template developed by André Miede.\footnote{\url{https://bitbucket.org/amiede/classicthesis/}}
The style was inspired by Robert Bringhurst's seminal book on typography
\emph{The Elements of Typographic Style}.\autocite{Bringhurst2004-wn}
The bibliography was processed using
Biblatex\footnote{\url{https://www.ctan.org/pkg/biblatex}} and
Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
Roberto Slimbach's Minion Pro provides both the text and display typeface.
\textsf{Open Sans} is the sans-serif typeface, while Jim Lyles's
\texttt{Bitstream Vera Mono} is used for monospaced text.
Finally, XITS acts at the math typeface.

\bigskip
{\small
\noindent\textit{\myTitle}
\\
\noindent\textcopyright\,\finalVersion, \myName
}

%\paragraph{note:} The custom size of the textblock was calculated
%using the directions given by Mr. Bringhurst (pages 26--29 and
%175/176). 10~pt Palatino needs  133.21~pt for the string
%``abcdefghijklmnopqrstuvwxyz''. This yields a good line length between
%24--26~pc (288--312~pt). Using a ``\emph{double square textblock}''
%with a 1:2 ratio this results in a textblock of 312:624~pt (which
%includes the headline in this design). A good alternative would be the
%``\emph{golden section textblock}'' with a ratio of 1:1.62, here
%312:505.44~pt. For comparison, \texttt{DIV9} of the \texttt{typearea}
%package results in a line length of 389~pt (32.4~pc), which is by far
%too long. However, this information will only be of interest for
%hardcore pseudo-typographers like me.%
%
%To make your own calculations, use the following commands and look up
%the corresponding lengths in the book:
%\begin{verbatim}
%    \settowidth{\abcd}{abcdefghijklmnopqrstuvwxyz}
%    \the\abcd\ % prints the value of the length
%\end{verbatim}
%Please see the file \texttt{classicthesis.sty} for some precalculated
%values for Palatino and Minion.
%
%    \settowidth{\abcd}{abcdefghijklmnopqrstuvwxyz}
%    \the\abcd\ % prints the value of the length