### 2016-10-03 FrontBackmatter/Colophon.tex
Finally get openright, right

\pagestyle{empty}

\hfill

\vfill

\pdfbookmark[1]{Colophon}{colophon}
\section*{Colophon}
This document was typeset using \LuaLaTeX, with a modified version of
the \texttt{classicthesis} template developed by André
Miede.\footnote{\url{https://bitbucket.org/amiede/classicthesis/}}
The style was inspired by Robert Bringhurst's seminal book on typography
\emph{The Elements of Typographic Style}\autocite{Bringhurst2004-wn}
and Edward Tufte's work.\autocite{Tufte1983-pr, Tufte1994-wf,
Tufte1997-fx, Tufte2006-ui}
The bibliography was processed using
Bib\LaTeX\footnote{\url{https://www.ctan.org/pkg/biblatex}} and
Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
TeX Gyre Termes provides both the text and display typeface.
\textsf{TeX Gyre Heros} is the sans-serif typeface, while Jim Lyles's
\texttt{Bitstream Vera Mono} is used for monospaced text.
Finally, XITS is the typeface for mathematical formulas.

\bigskip
{\small
\noindent\textit{\myTitle}
\\
\noindent\textcopyright\,\finalVersion, \myName
}