### 2016-10-03 Chapters/ResponseTheory.tex
Always link acronyms back to list of acronyms

\chapter{Response Theory and Molecular Properties}\label{ch:molprop}

\begin{epigraphs}
\qitem{
White light goin' messin' up my brain

White light it's gonna drive me insane
}{
--- \textsc{The Velvet Underground}
}
\qitem{What's the frequency, Kenneth? [...]

          I never understood the frequency, uh-huh}{
          --- \textsc{R.E.M.}}
\end{epigraphs}

Response theory provides an \emph{ab initio} framework for the
formulation and computation of molecular properties and can be
considered the connection between quantum chemistry and experimental
chemistry.
The response treatment of molecular properties has its roots in
time-dependent perturbation theory\autocite{Konishi2009-zb} and has been
continually developed in quantum chemistry for the past 30
years.\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}

Section \ref{sec:exact-response} will give a brief introduction to
response theory for isolated molecules.

\noindent The concepts of quasienergy,\autocite{Christiansen1998-pe} variational
perturbation theory\autocite{Helgaker1992-ph} and pole-and-residue
analysis of response functions\autocite{Olsen1985-nr} will be
introduced. The exposition closely follows the one by
\citeauthor{Saue2002-ns} in \noparcite[ref.][]{Saue2002-ns}.
I will explicitly derive the \acrshort{SCF} response function for a
quantum/classical polarizable Hamiltonian in Section
\ref{sec:csm-response}. The derivation will employ the
density matrix-based, \acrshort{AO} formalism introduced by
\citeauthor{Thorvaldsen2008-sg} in \noparcite[ref.][]{Thorvaldsen2008-sg}.
This formalism is suitable for the derivation of arbitrary-order
response functions and their residues, also when perturbation-dependent
basis sets are considered.\autocite{Friese2015-kb, Friese2015-bu}
\paper{V} presents the arbitrary-order derivation of the \acrshort{PCM}.

\section{Response Theory in a Nutshell}\label{sec:exact-response}

Let us consider the experimentally relevant situation where a molecular
system is subject to an external source of electromagnetic radiation,
Figure \ref{fig:OR}.
Here and in the following, "external" will be used to mean an
electromagnetic field which is "weak" when compared to the
electron--electron and electron--nuclei interactions in the molecule.

\begin{figure}[tb]
  \centering
  \includegraphics[width=.5\textwidth]{Optical-rotation.png}
  \caption{
  When a molecular system is subject to an external electromagnetic
  radiation of frequency $\omega$, a macroscopic, experimentally
  measurable response can be detected.
  Intensity and position of the signal can be related to the microscopic
  detail of the molecular system, as described by quantum mechanics.
  The picture depicts an optical rotation experiment where
  the linear polarization of the incident electromagnetic field
  is rotated by an angle $\alpha$ upon traversing a chiral material.
  Reproduced, with modifications, from \href{https://commons.wikimedia.org/wiki/File:Optical-rotation.svg}{Wikipedia}.
  }
  \label{fig:OR}
\end{figure}

The system can then be described by means of a combined matter--field
Hamiltonian:
\begin{equation}
  H = H_\mathrm{matter} + H_\mathrm{field} + H_\mathrm{int},
\end{equation}
where the three terms describe the molecular sample, the electromagnetic
field and their interaction, respectively.
Since our focus is on the molecular system, one can neglect the
quantized description of the field and treat the interaction term in a
semiclassical fashion.\autocite{Craig2012-zp}
The time-dependent Schr\"{o}dinger equation is now the appropriate
equation of motion:
\begin{equation}\label{eq:td-schrodinger}
  H\psi(t) = \mathrm{i}\pderiv{\psi(t)}{t},
\end{equation}
where the time-dependent, semiclassical matter--field Hamiltonian is
used:
\begin{equation}
  H = H_0 + V(t).
\end{equation}
The perturbation is a one-electron operator. We will further assume that
it is periodic in time $V(t) = V(t+T)$ and admits a discrete Fourier
decomposition:
\begin{equation}\label{eq:fourier}
  \begin{aligned}
 V(t) &=
 \sum_{k=-N}^N \expo{-\mathrm{i}\omega_k t}V(\omega_k)
 =
 \sum_{k=-N}^N\expo{-\mathrm{i}\omega_k t} \sum_X \epsilon_X(\omega_k)V_X \\
 &=
 \sum_{x}\expo{-\mathrm{i}\omega_x t}\epsilon_x V_X
  \end{aligned}
\end{equation}
where all the frequencies are integer multiples, $\omega_k=n_k \omega$
of the fundamental frequency $\omega = \frac{2\pi}{T}$.\autocite{Olsen1985-nr}
Frequencies and amplitudes in the Fourier decomposition obey the
following relations:
\begin{alignat}{2}
 \omega_{-k} = \omega_{k},
 \quad&
 \epsilon_x = \epsilon_X^*(\omega_k)
 =
 \epsilon_X(\omega_{-k}) = \epsilon^*_{-x},
\end{alignat}
since the $V(t)$ is required to be Hermitian.
For notational convenience, the perturbation operator ($X$) and
frequency ($k$) indices have been collapsed into the common index $x$.

Under the assumption that the external field is weak with respect to the
molecular field, time-dependent perturbation theory is an appropriate
tool to approximately solve Eq. \eqref{eq:td-schrodinger}.
The response theory route to molecular properties introduces a
phase-separated ansatz for the time-dependent wave function:
\begin{equation}\label{eq:phase-separated}
  \psi(t) = \expo{-\mathrm{i}F(t)}\bar{\psi}(t),
\end{equation}
where the time-derivative of the phase factor defines the time-dependent
\emph{quasienergy}:
\begin{equation}
  Q(t) = \dot{F}(t) = \Braket{\bar{\psi}(t) |
  H_0 + V(t) - \mathrm{i}\pderiv{}{t}
  | \bar{\psi}(t)}.
\end{equation}
The time-dependent quasienergy is variational, but the corresponding
time-dependent Hellmann--Feynman theorem cannot straightforwardly be
applied to the calculation of molecular properties.\autocite{Christiansen1998-pe, Pawlowski2015-sq}
Fortunately, we can exploit the periodicity of the perturbation operator
to introduce the time-averaged quasienergy:
\begin{equation}
  \aveQ =
 \frac{1}{T} \int_{-\frac{T}{2}}^{\frac{T}{2}} \diff t Q(t) =
 Q_0 + \sum_{x}\epsilon_x E_x
\end{equation}
where:
\begin{subequations}
 \begin{align}
  Q_0 &= \left\lbrace \Braket{\bar{\psi}(t) | H_0 | \bar{\psi}(t)} \right\rbrace_T
- \left\lbrace \Braket{\bar{\psi}(t) | \mathrm{i} \pderiv{}{t} | \bar{\psi}(t)}\right\rbrace_T = E_0(0) - S \\
  E_x &= \left\lbrace \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)}\exp(-\mathrm{i}\omega_k t) \right\rbrace_T
 \end{align}
\end{subequations}
Differentiating the time-averaged quasienergy yields the time-averaged
Hellmann--Feynman theorem:
\begin{equation}\label{eq:tave-hellfeyn}
  \deriv{\aveQ}{\epsilon_X(\omega_k)}
  =
  \left\lbrace\Braket{\bar{\psi}(t) | \pderiv{H}{\epsilon_x} | \bar{\psi}(t)}\right\rbrace_T
  = E_x,
\end{equation}
which can be used to connect response functions to molecular properties.
Letting $\ket{0}$ represent the unperturbed reference state, we can
develop the Kubo expansion at zero perturbation strength of the
expectation value of the observable $H_X$:\autocite{Kubo1957-ay}
\begin{equation}
\begin{aligned}
 \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)} &\simeq
 \Braket{0|V_X|0} +\sum_{k=-N}^N
 \response{V_X}{V(\omega_k)}{\omega_k}\expo{-\mathrm{i}\omega_k t} \\
 &+\frac{1}{2}\sum_{k,l=-N}^N
 \response{V_X}{V(\omega_k),V(\omega_l)}{\omega_k,\omega_l}\expo{-\mathrm{i}(\omega_k+\omega_l)t}
\end{aligned}
\end{equation}
response functions may be identified as the Fourier coefficients in the
expansion. These are none other but higher order derivatives of the
time-averaged quasienergy, as a comparison with the same expansion for
the right-hand side of Eq. \eqref{eq:tave-hellfeyn} will confirm:
\begin{equation}\label{eq:kubo}
\begin{aligned}
  \deriv{\aveQ}{\epsilon_x}
  &\simeq \Braket{0|V_X|0}\delta_{\omega_x} \\
  &+\sum_{y}\response{V_X}{V_Y}{\omega_y}\epsilon_y\delta_{\omega_x+\omega_y} \\
  &+\frac{1}{2}\sum_{y,z}\response{V_X}{V_Y,V_Z}{\omega_y,\omega_z}\epsilon_y\epsilon_z\delta_{\omega_x+\omega_y+\omega_z}
\end{aligned}
\end{equation}
First and higher order molecular properties can now be identified from
the response functions appearing as Fourier coefficients in the Kubo
expansion by taking the appropriate perturbation-strength derivatives at
zero field strength:
\begin{equation}\label{eq:nth-order-derivative}
  \left.\frac{\diff^n \aveQ}{\diff\epsilon_{x_1}\diff\epsilon_{x_2}\cdots\diff\epsilon_{x_n}} \right|_{\vect{\epsilon}=0}
  =
  \response{V_{X_1}}{V_{X_2}, \ldots, V_{X_n}}{\omega_{x_2},\ldots,\omega_{x_n}}\delta_{\omega_{x_1}+\cdots+\omega_{x_n}},
\end{equation}
the $\delta_{\omega_{x_1}+\cdots+\omega_{x_n}}$ notation is used to
enforce the sum rule on the probing and response frequencies involved in
the expansion:
\begin{equation}\label{eq:sum-rule}
 \sum_{i=1}^{n} \omega_{k_i} = 0
\end{equation}
Response functions quantify how, to a certain order in the perturbing
field tuples, the expectation value of a given observable is modified.
The semicolon appearing in the response functions separates the operator
$V_X$, whose response is measured, from the external perturbations
$V_Y$, $V_Z$\dots{} causing it.
From the perturbation expansion Eq. \eqref{eq:kubo}, it is clear that
permutation symmetry exists between the external perturbations provided
that the corresponding frequencies are permuted concomitantly.

Having introduced response functions as a useful tool to quantify the
effect of external perturbations on molecular properties, it is now time
to consider how to compute them.
In this Section, we are concerned with exact wave functions that can
be variationally parametrized by a set of parameters
$\vect{\lambda}$: $\ket{0}=\ket{0(\vect{\lambda})}$.
We recall that the Lagrangian method outlined in Section
\ref{sec:coupled-cluster} can be employed also in this
case.\autocite{Christiansen1998-pe, Helgaker2012-cz, Pawlowski2015-sq}
The optimal parameter set is determined upon minimization of the
time-averaged quasienergy:
\begin{equation}\label{eq:var-cond}
 \deriv{\aveQ}{\vect{\lambda}} = 0.
\end{equation}
In variational perturbation theory,\autocite{Helgaker1992-ph,
Saue2002-ns} we assume the variational parameters to be functions of the
perturbation strength parameters: $\vect{\lambda} =
\vect{\lambda}(\vect{\epsilon})$.
Furthermore, we require the stationarity condition to hold at \emph{all}
perturbation strengths, with $\lambda(0) = 0$, for convenience.

The perturbation-strength total derivative at zero perturbation strength
of the time-averaged quasi energy is:
\begin{equation}
\begin{aligned}
 \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]\aveQ
 &= Q^{[n]}_{x_1 \ldots x_n}
 =
 Q^{[n]}_{0; x_1 \ldots x_n}
 + \sum_{j=1}^{n}E^{[n-1]}_{x_j;x_1\ldots x_{j-1}x_{j+1}\ldots x_n} \\
 &=
 \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]Q_0
 + \sum_{j=1}^n
 \left[\prod_{i\neq j}^{n-1} \deriv{}{\epsilon_{x_i}}\right]E_{x_j}
\end{aligned}
\end{equation}
Notice that this is just an alternative and more verbose notation for
Eq. \eqref{eq:nth-order-derivative}, which allows to express response
functions in terms of wave function perturbed parameters.
Application of the chain rule to the total derivatives in fact yields:
\begin{subequations}
 \begin{align}
   Q^{[1]}_{a} &=
   \sum_{\sigma}\left[
   Q^{[1]}_{0; \sigma}\lambda^{[1]}_{\sigma;a} + E_{a}
   \right]\delta_{\omega_a} \\
   Q^{[2]}_{ab} &=
   \sum_{\sigma\tau}\left[
   \lambda^{[1]}_{\sigma;a}Q^{[2]}_{0;\sigma\tau}\lambda^{[1]}_{\tau;b}
   + Q^{[1]}_{0; \sigma}\lambda^{[2]}_{\sigma;a}
   + E^{[1]}_{a;\sigma}\lambda^{[1]}_{\sigma;b}
   + E^{[1]}_{b;\sigma}\lambda^{[1]}_{\sigma;a}
   \right]\delta_{\omega_a+\omega_b} \label{eq:linres}
 \end{align}
\end{subequations}
where the following tensors were introduced:
\begin{subequations}
 \begin{align}
 Q^{[n]}_{0; \sigma_1\cdots \sigma_n} &=
 \frac{\partial^n Q_0}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
 E^{[n]}_{x;\sigma_1\cdots \sigma_n} &=
 \frac{\partial^n E_x}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
\lambda^{[n]}_{\sigma;x_1\dots x_n} &=
\frac{\partial^n \lambda_\sigma}{\partial \epsilon_{x_1}\cdots\partial \epsilon_{x_n}}
 \end{align}
\end{subequations}

The response equations, determining the tensor of
perturbed wave function parameters $\vect{\lambda}^{[n]}$, are obtained by differentiating the
variational condition on the time-averaged quasienergy Eq.
\eqref{eq:var-cond}:\autocite{Olsen1985-nr, Helgaker1992-ph, Christiansen1998-pe}
\begin{equation}
 R_{\sigma;x_1\cdots x_n}^{[n]} = \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]
\left. \deriv{\aveQ}{\lambda_\sigma}\right|_{\vect{\epsilon}=0} = 0.
\end{equation}
The determination of second-order properties requires the knowledge of
the first-order response of the wave function:
\begin{subequations}
 \begin{align}
 R_{\sigma}^{[0]} &= Q_{0;\sigma}^{[1]} = 0 \\
 R_{\sigma;a}^{[1]} &= \sum_{\tau}Q_{0;\sigma\tau}^{[2]}\lambda_{\tau;a}^{[1]} + E_{a;\sigma}^{[1]} = 0.
 \label{eq:1st-order-response}
 \end{align}
\end{subequations}
The time-averaged quasienergy formalism allows to formulate response
properties to static and dynamic fields on an equal footing as
perturbation-strength derivatives.
Moreover, variational perturbation theory achieves a transparent
derivation of the necessary response equations.\autocite{Helgaker1992-ph}

Let us consider in more detail the form of the linear response function
Eq. \eqref{eq:linres}. The formal solution to the response equation Eq.
\eqref{eq:1st-order-response} can be obtained by inverting the
\emph{Hessian}. Eq. \eqref{eq:linres} can then be rewritten as:
\begin{equation}
  \response{V_A}{V_B}{\omega_b}
  =
   -\sum_{\sigma\tau}
   E^{[1]}_{a;\sigma}
   \left[\mat{Q}^{[2]}_{0}\right]^{-1}_{\sigma\tau}
   E^{[1]}_{b;\tau}.
\end{equation}
Expansion into the exact eigenbasis of $H_0$ yields the
\gls{SOS} expression for the linear response function:
\begin{equation}
  \response{V_A}{V_B}{\omega_b}
  =
  -\sum_{n>0}\left[
  \frac{\braket{0|V_A|n}\braket{n|V_B|0}}{\omega_{n0}-\omega_b}
  +
  \frac{\braket{0|V_B|n}\braket{n|V_A|0}}{\omega_{n0}-\omega_a}
  \right]\delta_{\omega_a+\omega_b}.
\end{equation}
The excitation energies of the system appear in the denominators
$\omega_{n0} = E_n - E_0$ and can thus be identified as the \emph{poles}
of the response function. Moreover, the \emph{residues} at
the poles are the transition moments determining the intensity of
the spectroscopic transitions:
\begin{equation}
  \lim_{\omega_b\rightarrow\omega_{n0}}
  \response{V_A}{V_B}{\omega_b}
  = \braket{0|V_A|n}\braket{n|V_B|0}.
\end{equation}
A similar analysis holds also for higher order response functions. The
pole structure of the higher order quantities also conveys information
about the excited states \emph{via} single and double
residues.\autocite{Olsen1985-nr, Christiansen1998-pe, Helgaker2012-cz}

The derivation above holds for an exact state, a situation that is never
realized in common practice. However, response theory is
also applicable to approximate states.
As noted by \citeauthor{Norman2011-ad}, a response theory treatment of
molecular properties can be summarized into the following four
steps:\autocite{Norman2011-ad}
\begin{enumerate}
    \item Single out a quantity of interest that can be connected to
      molecular properties. In our case, the time-averaged
      quasienergy;
    \item Introduce a suitable parametrization of the time-dependent,
      phase-separated wave function;
    \item Devise the appropriate equation of motion, based on the
      time-dependent Schr\"{o}dinger equation, for the time-dependent
      wave function parameters;
    \item Identify response functions and response equations.
\end{enumerate}
Thus, one can formulate and compute molecular properties based on
the approximate state ans\"{a}tze of quantum chemistry.
An interesting formulation of \acrshort{SCF} response theory is the
open-ended \acrshort{AO} density matrix-based approach presented by
\citeauthor{Thorvaldsen2008-sg}.\autocite{Thorvaldsen2008-sg}
By formulating the perturbation-strenght derivative of the time-averaged
quasienergy Lagrangian in terms of the \acrshort{AO} density matrix,
the authors were able to formulate arbitrary order response functions
in the \acrshort{AO} basis.
The density matrix is the variational degree of freedom and the response
equations can be obtained by differentiation of the \gls{TDSCF}
equation, the stationarity condition on the density matrix.
The approach lends itself to an efficient \acrshort{AO}-based
implementation,\autocite{Larsen2000-hj, Kjaergaard2008-hy} can fully
leverage different response parameters elimination
schemes,\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
to an open-ended, recursive implementation.\autocite{Ringholm2014-gx,
Friese2015-kb, Friese2015-bu}
We will base our derivation of \acrshort{SCF} response theory for
quantum/classical polarizable Hamiltonians of this framework.

\section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}

The development of response theory for quantum/classical polarizable
Hamiltonians has moved almost at the same pace as the polarizable
models themselves.
An ample literature exists in the context of the
\acrshort{PCM},\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
polarizable \acrshort{MM} models,\autocite{Curutchet2009-bt, Olsen2010-wa,
Lipparini2012-hx, Lipparini2012-tl}
and general \acrshort{QM}/\acrshort{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}

Solvent relaxation is an important point when formulating response
properties. Solvation dynamics occurs on timescales and through
relaxation modes that are typical of the solvent molecular structure and of the
peculiar solute-solvent interactions.
This is clearly difficult to model with continuum models, since the
entire frequency spectrum of the permittivity $\diel(\omega)$ would be
required.\autocite{Ingrosso2003-ev, Caricato2005-lo,
Mennucci2005-vi, Caricato2006-ba, Corni2015-pe}
However, if one considers the initial steps of an electronic excitation
process, a Franck--Condon-like response of the solvent can be assumed.
Nuclear motions of the solute and solvent molecules alike will not be able to
follow the fast rearrangement of the electronic density, therefore the
corresponding part of the response will remain frozen in the state
immediately preceding the change.
Such an assumption, albeit approximate, leads to the following partition
of the polarization vector:
The polarization vector will then be (approximately) a sum of two
components:
\begin{equation}
 \vect{P} \simeq \vect{P}_\mathrm{fast} + \vect{P}_\mathrm{slow}.
\end{equation}
This is equivalent to considering only the asymptotic limits of the solvent
dielectric spectrum, \ie the static $\diels$ and optical
dielectric constants $\dield$.\autocite{Aguilar1993-pu, Cammi1995-ah,
Cammi2005-tb, Corni2005-nd}

The variational setting introduced in Section \ref{sec:variational},
offers a series of advantages also in the formulation of response
theory. As already noted, a classical counterpart to the
Hellmann--Feynman theorem holds:
\begin{equation}\tag{\ref{eq:classical-HF} from Chapter \ref{ch:CSM}}
  \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
  + \pderiv{U_\mathrm{pol}}{\s}\pderiv{\s}{F}
  = \pderiv{U_\mathrm{pol}}{F},
\end{equation}
which forms the basis for our formal development in
\paper{V}.\autocite{pcm-openrsp}
There we show how the variational polarization functional can
successfully achieve a transparent derivation of arbitrary order
response functions, when the fixed-cavity approximation for the
\acrshort{PCM} is assumed.
Here we report a summary of our findings in \paper{V} employing the
supermatrix form of the classical functional as given in Eq.
\eqref{eq:supermatrix-functional}.
In the notation of \citeauthor{Thorvaldsen2008-sg}, we introduce the
generalized free energy:
\begin{equation}\label{eq:generalized-KS-energy}
  \begin{aligned}
 \tilde{\mathcal{G}}
 &=
 \tilde{\mathcal{E}}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p} +
 {}^t\tilde{\p}\Tr(\tilde{\mathbfup{s}}\pertmat{D}) \\
 &\eqtr
 \left[\pertmat{h} + \pertmat{V}^{t} +
 \frac{1}{2}\pertmat{G}^{\gamma}(\pertmat{D}) -
 \frac{\mathrm{i}}{2}\pertmat{T}\right]\pertmat{D}
 + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
 + {}^t\tilde{\p}\tilde{\mathbfup{s}}\pertmat{D},
 \end{aligned}
\end{equation}
whose perturbation-strength derivatives at zero perturbation strength
will yield the desired response functions.

To keep track of the number of chain rule applications,
we will denote higher-order derivatives of the \acrshort{KS}
generalized free energy with the composite ${mn, abc\ldots}$
superscript:
\begin{equation}\label{eq:differential}
  \mathcal{G}^{mn, abc}
  =
  \frac{\partial^{m+n+3}\mathcal{G}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}
  =
    \mathcal{E}^{m, abc} +
  \frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c} \text{.}
\end{equation}
In this notation, the $abc\ldots$ indices denote the perturbations with
respect to which differentiation occurs. The index $m$ denotes the order
of differentiation with respect to the density matrix. $n$
denotes the order of differentiation with respect to the
polarization.
The classical interaction kernel $\V$ might, in the most general
scenario, depend on the molecular geometry.
In \paper{V} and in the following we ignore this possibility and work
under the simplifying assumption that $\V$ is independent of the nuclear
positions.
This goes under the name of \emph{fixed-cavity approximation} in the
context of the \acrshort{PCM}.\autocite{Cammi1994-qj}
Under this simplifying assumption, the perturbation-strength derivative
of the classical functional will only include terms stemming from the
linear interaction term:
\begin{equation}\label{eq:polarization-derivative}
 \deriv{}{\epsilon_a}\left\lbrace
 U_\mathrm{pol}
 \right\rbrace_T
 \eqavetr
 {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D} \text{,}
\end{equation}
where the second term only involves derivatives of the source integrals.
Moreover, in the general expression for the perturbation-strength
derivative of the polarization functional in Eq. \eqref{eq:differential},
the $m$ and $n$ indices can only assume the values 0 or 1.
In fact, the density matrix dependence is at most linear in the
polarization functional, while the classical Hellmann--Feynman theorem
ensures that the polarization appears at most linearly in $\tilde{\mathcal{G}}^{00,a}$:
\begin{equation}\label{eq:tilde-G00-a}
  \begin{aligned}
    \tilde{\mathcal{G}}^{00,a}(\tilde{\mat{D}}, \tilde{\sigma}, t)
    &= \tilde{\mathcal{E}}^{0,a} +
    {}^t\tilde{\p}\Tr({\tilde{\mathbfup{s}}^a\pertmat{D}}) \\
  &\eqtr
  (\pertmat{h}^a + \pertmat{V}^{t,a} + \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
  \end{aligned}
\end{equation}
$\pertmat{F}^{\Omega_a}_\mathrm{xc}$ is the functional derivative matrix
defined in terms of the perturbed overlap distributions
$\tilde{\Omega}^a$.
Further perturbation-strength differentiations and evaluation at zero
perturbation strength will yield the desired response functions:
\begin{subequations}
  \begin{align}
    L^a &\eqavetr
    \mathcal{G}^{00,a} - \mat{S}^a\mat{W} \label{eq:first-order} \\
    L^{ab} &\eqavetr \mathcal{G}^{00,ab}
    + \mathcal{G}^{10,a}\mat{D}^{b}
    + \mathcal{G}^{01,a}\p^{b}
    - \mat{S}^{ab}\mat{W}
    - \mat{S}^{a}\mat{W}^{b} \label{eq:second-order} \\
    L^{abc} &\eqavetr
      \mathcal{G}^{00,abc}
    + \mathcal{G}^{10,ac}\mat{D}^{b}
    + \mathcal{G}^{10,ab}\mat{D}^{c}
    + \mathcal{G}^{20,a}\mat{D}^{b}\mat{D}^{c} \nonumber \\
    &+ \mathcal{G}^{10,a}\mat{D}^{bc}
    + \mathcal{G}^{11,a}\mat{D}^{b}\p^c \label{eq:third-order} \\
    &+ \mathcal{G}^{01,ac}\p^{b}
    + \mathcal{G}^{01,ab}\p^{c}
    + \mathcal{G}^{01,a}\p^{bc}
    + \mathcal{G}^{11,a}\p^{b}\mat{D}^c \nonumber \\
    &- \mat{S}^{abc}\mat{W}
    - \mat{S}^{ab}\mat{W}^{c}
    - \mat{S}^{ac}\mat{W}^{b}
    - \mat{S}^{a}\mat{W}^{bc} \nonumber
  \end{align}
\end{subequations}
and similarly for higher orders.

The governing equations for the perturbed polarization densities are
obtained in complete analogy with the handling of pertubed density
matrices described in \noparcite[ref.][]{Thorvaldsen2008-sg}
Denoting by the superscript $b_N$ a tuple of simulatenous perturbations,
the polarization can be decomposed into frequency components.
Similarly to the treatment of the perturbed density matrices,
we introduce a partition of the frequency components of the
polarization $\p_\omega^{b_N}$ into homogeneous (H) and particular (P)
terms:
\begin{equation}\label{eq:polarization-partition}
  \p_\omega^{b_N} = \p_\mathrm{P}^{b_N} + \p_\mathrm{H}^{b_N} \text{,}
\end{equation}
and we thus get the following system of equations:
\begin{subequations}
\begin{empheq}[left={\empheqlbrace}]{align}
    \V\p_\mathrm{H}^{b_N} &+ \Tr \mathbfup{s}\Dh{b_N} = 0
    \label{eq:homo-polarization-eq}\\
    \V\p_\mathrm{P}^{b_N} &+ \Tr \mathbfup{s}\Dp{b_N}
    = \mathbb{S}^{(n-1)}_\omega \text{.}
 \label{eq:part-polarization-eq}
\end{empheq}
\end{subequations}
$\mathbb{S}^{(n-1)}_\omega$ contains all terms that depend on lower
order density matrices and differentiated source integrals.
The term $\mathbb{S}^{(n-1)}_\omega$ always contains at least a first
derivative of the source integrals and is thus zero if the basis set is
independent of the perturbation tuple under consideration.
Consequently, the particular term of the polarization is nonzero if and
only if the basis set depends on the external perturbation.

Response equations can be obtained by differentiating the
\acrshort{TDSCF} equation with respect to the perturbation tuple
$b_N$:\autocite{Thorvaldsen2008-sg, Ringholm2014-gx}
\begin{equation}\label{eq:response-equation}
 \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} = \RHS{b_N}
\end{equation}
where the generalized Hessian $\genHessian$ and metric $\genMetric$
matrices were introduced and are defined by their transformations on the
response parameters $\rspParam{b_N}$:\autocite{Larsen2000-hj,
Kjaergaard2008-hy}
\begin{subequations}
\begin{align}
  \genHessian\rspParam{b_N} &=
  \mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})\mat{D}\mat{S}
  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N},
  \mat{D}]_{\mat{S}}) \nonumber \\
&+\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
-\mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{F}
\label{eq:E2-lintra} \\
&+ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
-\mat{S}\mat{D}\mat{\esp}\ASCh{b_N} \nonumber
\\
  \genMetric\rspParam{b_N} &= \mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S} \text{.}
  \label{eq:S2-lintra}
\end{align}
\end{subequations}
The generalized Hessian matrix $\genHessian$ includes two types of
classical terms: implicit and explicit. \emph{Implicit} terms are those
included in the zeroth-order Fock matrix, $\mat{F}$.
The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
as they explicitly involve the $N$-th order homogenous polarization
variational density.
The nonequilibrium response framework needed in continuum models when
dealing with frequency-dependent properties,\autocite{Cammi1999-rb, Tomasi2005-vm}
will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
evaluated by making use of the optical permittivity $\dield$ in the
polarization equations \eqref{eq:part-polarization-eq}.
The \acrlong{RHS} in the response equation only includes terms that
depend on particular contributions up to the desired order or
lower-order perturbed variational densities:\autocite{Ringholm2014-gx}
\begin{equation}\label{eq:RHS}
  \RHS{b_N} =
  \left[\left( \tilde{\mathbfcal{F}}
-
\frac{\mathrm{i}}{2}\pertmat{S}\deriv{}{t}\right)\pertmat{D}\pertmat{S}\right]^{\ominus,
b_N}_\mathrm{P},
\end{equation}
where $\mathbfcal{F}$ is the generalized \acrshort{KS} matrix, see
\noparcite[ref.][]{Thorvaldsen2008-sg} for details.

A direct solution of the response equation \eqref{eq:response-equation} by inversion
is not feasible due to the size of the $\left[\genHessian -
\omega_{b_N}\genMetric \right]$ matrix. \emph{Subspace iteration}
methods are usually employed and the solution is achieved iteratively
by enlarging a projection subspace of trial vectors.\autocite{Saad2003-oa,
Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}
These methods can be implemented efficiently and in matrix-free form
since the linear transformations $\genHessian\rspParam{b_N}$ and
$\genMetric\rspParam{b_N}$ have known forms only involving integral and
(perturbed) density matrices.
Moreover, Jacobi precondinioting of the iterations is usually sufficiently
robust and efficient, due to the diagonal dominance of the generalized
Hessian.