### 2016-10-04 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\chapter*{Acknowledgments}

It has not been one day, it's been four years.\footnotetext{Just a perfect day \\
You made me forget myself \\
I thought I was \\
Someone else, someone good \\
\hspace*{15pt}
--- \textsc{Lou Reed}}
And it's not always been perfect.
It is however true that the people I have met and interacted with during
this adventure have made me feel \emph{someone good}.
I don't feel it's an exaggeration to say that if I have made it this far
is also because of \emph{you}.
Thus, these few, poorly written lines are
a \emph{dedication} of this dissertation to you
and an \emph{acknowledgment}
of the patience and love you have shown towards me.
Thanks for not giving up on me.

First and foremost, \textitalian{ai miei genitori,
Adriano e Vilma. Ai miei fratelli, Francesca e Simone. A mia zia, Marina
ed ai miei nonni, Cherubino e Irma, Maria e Piero. Questo è dedicato a
voi, per tutto quello che mi avete insegnato, per il sostegno continuo e
la pazienza infinita che dimostrate ogni giorno.}

in no particular order:\footnotetext{And all of my friends were there, \\
and no one cared \\
\hspace*{15pt}
--- \textsc{The Kinks}}
\begin{itemize}
  \item To \emph{Radovan Bast}, \emph{il miglior fabbro},
  \item To \emph{Davide Michetti}, the inventor of nicknames,
  \item To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and
    Torgeir,
  \item To the \emph{Schrodinger killed my cat!}
\end{itemize}

\textitalian{A Grazia, Graziella e al Toblerone, qualcuno sa perché.}