### 2016-10-04 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\chapter*{Acknowledgments}

\epigraph{Just a perfect day

You made me forget myself

I thought I was

Someone else, someone good
}{
--- \textsc{Lou Reed}}

It has not been one day, it's been four years.
And it's not always been perfect.
It is however true that the people I have met and interacted with during
this adventure have made me feel \emph{someone good}.
I don't feel it's an exaggeration to say that if I have made it this far
is also because of \emph{you}.
Thus, these few, poorly written lines are
a \emph{dedication} of this dissertation to you
and an \emph{acknowledgment}
of the patience and love you have shown towards me.
Thanks for not giving up on me.

First and foremost, \textitalian{ai miei genitori,
Adriano e Vilma. Ai miei fratelli, Francesca e Simone. A mia zia, Marina
ed ai miei nonni, Cherubino e Irma, Maria e Piero. Questo è dedicato a
voi, per tutto quello che mi avete insegnato, per il sostegno continuo e
la pazienza infinita che dimostrate ogni giorno.}
To my advisor, Luca Frediani, for leaving me ample freedom of exploring
and experimenting in these past four years.
It's been fun working with you. You have made me feel like a part of
your family and this is also for Emil and Emma.
To my co-advisor, Benedetta Mennucci, for always taking the time to
answer my questions.
To my friends from high school, for the good times we've had and for
those we will have.
To the \enquote{Pisa diaspora} and in particular to Stefano, Federica,
Laura and Marco. Remember, this is almost your last chance to visit me in Tromsø
before I disinherit you!
A special mention to Simone, my numerical analysis guru.
To the \enquote{Repubblica di Cisanello}: Giacomo, Emilio and Romualdo.
For the fun and conversations we've had. And for the, now legendary,
late-night couch moving session across town with Daniele and Matteo.
To all the MoLEcoLabbers, thank you for the great environment and
hospitality whenever I visit Pisa.
To all the members of CTCC, for the stimulating environment, the
interesting discussions, the excellent cakes and all the wine bottles I
have won.
To Radovan, \emph{il miglior fabbro}, for all I have learnt
from you.
To my flatmates, Davide, Thibaud, Eivind, Audun, Jenn and JB. For making
my daily life in Tromsø less lonely and more fun.
Ah, yeah, sorry for the loud blasts of noise I like to listen to!
To Alex, that always has a word of advice and encouragement.
To Maarten and Anna, for their precious friendship and for pushing me to
speak Norwegian more.
To Laura, the Finnish lady who thought it would be fun to bring me along
on a camping-and-bouldering trip to Lofoten. And it actually was!
Despite the rain. And me not actually bouldering. But that's another
story.
To the Italian \enquote{colony} in Tromsø, thank you. Tromsø feels like
home also because of you.
To Laura, for not giving up on me, despite the evenings spent at the
filmklubb.
To Lindon, the coolest organic chemist. Thanks for all the music tips
and the amazing time at the Rock Your Head Festival.
To Glenn, for being an awesome friend.
To Ravna and Lili, for your kindness and the hours spent at the swimming
pool together.
To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and Torgeir.
One day we'll leave \emph{The Monroeville Zombies} in the dust, I am
sure.
To T. Daniel Crawford, for his hospitality, enthusiasm and patience with
my slow progress.
My visits to Blacksburg would not have been possible without the warm
hospitality of Laurie, Charles and Alec. This is dedicated to you too!
To Marco, because the world really is small and finding family where you
least expect it is always a great surprise.
To Yang, Töph, Pablo, Tommaso and Lorenzo, for the great fun we've had at the
schools.

\textitalian{A Grazia, Graziella e al Toblerone, qualcuno sa perché.}