### 2016-12-05 FrontBackmatter/Errata.tex
Add Errata Corrige to the submitted and approved version of the dissertation

\pagestyle{empty}

\pdfbookmark[1]{Errata Corrige}{errata}
\chapter*{Errata Corrige}

\begin{itemize}
\item On page 23. The labelling of the cavity boundary in Figure 2.1 is
inconsistent. \enquote{$\Gamma\equiv\Omega_\mathrm{i}$} should be replaced by
\enquote{$\Gamma\equiv\partial\Omega_\mathrm{i}$}, as in the following Figure.
  \begin{figure*}[!h]
    \centering
    \includegraphics[width=.5\textwidth]{IEF-correct.pdf}
  \end{figure*}
\item On page viii, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page 96, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page 210, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page vii, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 94, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 186, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 126, there is a misprint in the title for reference 225.
\enquote{RNa} and \enquote{DNa} should be replaced by \enquote{RNA} and
\enquote{DNA}, respectively.
\item On page viii, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item On page 96, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item On page 210, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item Paper IV was accepted for publication on September 16, 2016. The paper
was published in final form on the journal's website shortly thereafter. The
\emph{verbatim} copy included in the dissertation reflects the accepted version
of Paper IV.
\item Paper V was accepted for publication on November 15, 2016. The paper was
published in final form on the journal's website shortly thereafter. The
\emph{verbatim} copy included in the dissertation reflects the submitted
version of Paper V, which underwent minor revisions after the peer-review
process.
\end{itemize}