### 2017-01-27 FrontBackmatter/Publications.tex
This is the final printed version

%*******************************************************
% Publications
%*******************************************************
\thispagestyle{empty}
\phantomsection
\refstepcounter{dummy}
\addcontentsline{toc}{chapter}{\tocEntry{Publications}}

\chapter*{Publications}

This thesis is based on the following scientific papers:

\renewcommand{\descriptionlabel}[1]{\hspace*{\labelsep}#1}
\begin{description}[leftmargin=2mm, style=nextline]
\item[\paper{I}]
  {\small\textsf{
  Four-Component Relativistic Calculations in Solution with the
  Polarizable Continuum Model of Solvation: Theory,
  Implementation, and Application to the Group 16 Dihydrides
  \ce{H2X} (\ce{X} = \ce{O}, \ce{S}, \ce{Se}, \ce{Te},
  \ce{Po})
  }
  \\
  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
  \\
\textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
  \\
  DOI: \url{10.1063/1.4943782}
  }
\label{relapcm}

\item[\paper{II}]
  {\small\textsf{
  Wavelet Formulation of the Polarizable Continuum Model. II. Use of
  Piecewise Bilinear Boundary Elements
  }
  \\
  M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
  Harbrecht,  and L. Frediani
  \\
  \textit{Phys. Chem. Chem. Phys.}, \textrm{2015}, \textbf{17},
  31566--31581
  \\
  DOI: \url{10.1039/C5CP03410H}
  }
\label{wemlin}

\item[\paper{III}]
  \small{\textsf{
  A Polarizable Continuum Model for Molecules at Spherical
  Diffuse Interfaces
  }
  \\
  \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
  Frediani
  \\
  \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103
  \\
  DOI: \url{10.1063/1.4943782}
  }
\label{spherical}

\item[\paper{IV}]
  \small{\textsf{
  Four-Component Relativistic Density Functional Theory with the
  Polarisable Continuum Model: Application to EPR Parameters
  and Paramagnetic NMR Shifts
  }
  \\
  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
  Frediani, and K. Ruud
  \\
  Accepted for publication in \textit{Mol. Phys.}
  \\
  DOI: \url{10.1080/00268976.2016.1239846}
  }
\label{pcmepr}

\item[\paper{V}]
  {\small\textsf{
  Open-Ended Formulation of Self-Consistent Field Response Theory with
  the Polarizable Continuum Model for Solvation
  }
  \\
  \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
  A. H. Steindal, K. Ruud, and L. Frediani
  \\
  Submitted to \textit{Phys. Chem. Chem. Phys.}
  }
\label{pcmopenrsp}
\end{description}
\renewcommand{\descriptionlabel}[1]{\hspace*{\labelsep}\spacedlowsmallcaps{#1}}

\section*{Notes Added in Proof}

\begin{itemize}
\item \paper{IV} was published online in final form on October 18, 2016.
  The \emph{verbatim} copy included in the dissertation reflects the
  accepted version of \paper{IV}.
\item \paper{V} was accepted for publication on November 15, 2016 and
  assigned the DOI \url{10.1039/C6CP06814F}.
  The paper was published online in final form on December 1, 2016.
  The \emph{verbatim} copy included in the dissertation reflects the
  submitted version of \paper{V}, which underwent minor revisions after
  the peer-review process.
\end{itemize}